---
title: "datasaurus test"
date: "2025-11-20"
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

# Demo DatasauRus Dozen Dataset

## Intro

🦕 The Datasaurus Dozen data:  

The package contains 13 different datasets, one of which is the famous Dino dataset (the "Datasaurus" shape).  

Same Statistics:  

All 13 datasets in the "Datasaurus Dozen" have nearly identical or exactly the same basic summary statistics, such as the mean of $x$ and $y$ variables, the standard deviation of $x$ and $y$, and the correlation between $x$ and $y$.

Different Visualizations:  
When plotted, these datasets form radically different visual shapes, including a star, a circle, and the iconic dinosaur.  


Why it is important:  

>By providing data that is statistically indistinguishable but visually distinct, the datasauRus package drives home the point that plotting your data is crucial for understanding its underlying structure.

```{r}
library(tidyverse)
library(datasauRus)
```


```{r}
head(datasaurus_dozen)
```


```{r}
datasaurus_dozen %>%
    group_by(dataset) %>%
    summarize(
      mean_x    = mean(x),
      mean_y    = mean(y),
      std_dev_x = sd(x),
      std_dev_y = sd(y),
      corr_x_y  = cor(x, y)
    )
```
## Away

```{r}
datasaurus_dozen %>%
  filter(dataset == "away") %>%
  ggplot(aes(x = x, y = y)) +
    geom_point() +
    theme_void() +
    theme(legend.position = "none") 
```

## Dino

```{r}
datasaurus_dozen %>%
  filter(dataset == "dino") %>%
  ggplot(aes(x = x, y = y)) +
    geom_point() +
    theme_void() +
    theme(legend.position = "none") 
```

## Star

```{r}
datasaurus_dozen %>%
  filter(dataset == "star") %>%
  ggplot(aes(x = x, y = y)) +
    geom_point() +
    theme_void() +
    theme(legend.position = "none") 
```


## alles met facet wrap

```{r fig.width=15, fig.height=15}
ggplot(datasaurus_dozen, aes(x = x, y = y, colour = dataset)) +
    geom_point() +
    theme_void() +
    theme(legend.position = "none") +
    facet_wrap(~dataset, ncol = 3)
```


The End...

---